---
title: "Advert_Prediction"
author: "Joy Muli"
date: "11/4/2020"
output: 
  md_document:
    variant: markdown_github
---
```{r}
library(tidyverse)
```

# Reading the dataset
```{r}
adverts <- read.csv("http://bit.ly/IPAdvertisingData")
```
# Previwing the dataset
```{r}
head(adverts)
```
# Checking the bottom
```{r}
tail(adverts)
```
# Checking the number of rows and columns
```{r}
#The rows and columns in the data 
cat("The dataset has ", dim(adverts)[1], "rows and ", dim(adverts)[2], " columns")
```
# Checking the structure of the dataset
```{r}
str(adverts)
```

## Cleaning the dataset
# Making all the columns lower cased for uniformity and replacing white spaces
```{r}
names(adverts) <- str_replace_all(names(adverts), c(" " = "_"))
names(adverts) <- tolower(names(adverts))
```
# Identifying numeric columns
```{r}
num_cols <- unlist(lapply(adverts, is.numeric))
num_cols

```
# They are; daily time spent on site, age, area income, daily internet usage, male, click on ad
 # Evaluating if there are any outliers in the numeric class
#Subset numeric columns of data
```{r}
data_num <- adverts[ ,num_cols]                        
```
# Plotting boxplots
```{r}
boxplot(data_num)

```
# There were outliers only on the area_income. Now to check the values
```{r}
boxplot.stats(adverts$area.income)$out
```
# Dealing with the outliers 
# Capping
```{r}
x <- adverts$area.income
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
adverts$area.income[adverts$area.income < (qnt[1] - H)] <- caps[1]
adverts$area.income[adverts$area.income > (qnt[2] + H)] <- caps[2]
```
# Checking if there are any outliers left
```{r}
boxplot.stats(adverts$area.income)$out
```
# Checking for missing values
```{r}
colSums(is.na(adverts))

```
# There are no missing values
# Checking for duplicated columns
```{r}
duplicate_rows <- adverts[duplicated(adverts),]
duplicate_rows

```
# There are no duplicated columns


# Exploratory Analysis

## Measures of central tendency
# Getting the mean for each numeric column
```{r}
colMeans(data_num)
```
# Getting the median for each numeric column
```{r}
apply(data_num,2,median)

```
# Getting the mode
```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```
```{r}
getmode(adverts$age)
getmode(adverts$daily.time.spent.on.site)
getmode(adverts$area.income)
getmode(adverts$daily.internet.usage)
getmode(adverts$city)
getmode(adverts$country)
getmode(adverts$clicked.on.ad)
getmode(adverts$timestamp)
getmode(adverts$male)
```

## Measures of dispersion

# Checking the minimum for every numeric column
```{r}
adverts.min <- apply(subset(data_num, select = c(daily.time.spent.on.site, age, area.income,	daily.internet.usage)),2,min, na.rm = TRUE)
adverts.min
```
# Checking the maximum for every numeric column
```{r}
adverts.max <- apply(subset(data_num, select = c(daily.time.spent.on.site, age, area.income,	daily.internet.usage)),2,max, na.rm = TRUE)
adverts.max

```
# Checking the range for every numeric column
```{r}
adverts.range <- apply(subset(data_num, select = c(daily.time.spent.on.site, age, area.income,	daily.internet.usage)),2,range, na.rm = TRUE)
adverts.range
```
# Checking the quantiles for every numeric column
```{r}
adverts.qnt <- apply(subset(data_num, select = c(daily.time.spent.on.site, age, area.income,	daily.internet.usage)),2,quantile, na.rm = TRUE)
adverts.qnt

```
# Checking the variation for every numeric column
```{r}
adverts.var <- apply(subset(data_num, select = c(daily.time.spent.on.site, age, area.income,	daily.internet.usage)),2,var, na.rm = TRUE)
adverts.var
```
# Checking the standard deviation for every numeric column
```{r}
adverts.sd <- apply(subset(data_num, select = c(daily.time.spent.on.site, age, area.income,	daily.internet.usage)),2,sd, na.rm = TRUE)
adverts.sd
```

## Graphical Univariate
# Frequency distribution
```{r}
# Fetching the age column
age <- adverts$age

# Applying the table() function will compute the frequency distribution of the age variable
age_frequency <- table(age)

# Then applying the barplot function to produce its bar graph
barplot(age_frequency)
```
# Histogram to show the distribution of time spent daily on site
```{r}
hist(adverts$daily.time.spent.on.site)
```
# Distribution of age
```{r}
hist(adverts$age)
```
# area_income distribution
```{r}
hist(adverts$area.income, col='cadetblue')
```
# Distribution of daily internet usage
```{r}
hist(adverts$daily.internet.usage, col='cadetblue')

```
# Distribution of gender
```{r}

ggplot(adverts,aes(x=toupper(male)))+geom_bar()+xlab(label = "Males and Females")+ylab(label = "Frequency")+theme_classic()

```
# Distribution of clicked on ad
```{r}
ggplot(adverts,aes(x=toupper(clicked.on.ad)))+geom_bar()+xlab(label = "No or Yes")+ylab(label = "Frequency")+theme_classic()

```

# Summary
```{r}
summary(adverts)

```
# Checking the number of people who clicked on ads and who did not
```{r}
table(adverts$clicked.on.ad)
```
# Checking the number of men and women
```{r}
table(adverts$male)
```
# From this we can see that the dataset was pretty much balanced


## Bivariate Analysis

# Correlation
# Correlation matrix
```{r}
M<-cor(data_num)
head(round(M,2))

```
# Covariance
```{r}
# Finding the covariance between daily_time_spent_on_site and age
# Assigning a variable to each column
site_time <- adverts$daily.time.spent.on.site
age <- adverts$age
cov(site_time, age)
```

```{r}
# Covariance between daily_time_spent_on_site and income
income <- adverts$area.income
cov(site_time, income)
```

# Scatter plot
```{r}
# Scatter plot to compare income vs time spent on site
plot(site_time, income, xlab="Time spent on site", ylab="Area income")
```
```{r}
clicked <- adverts$clicked.on.ad
# Scatter plot to compare income vs clicked on ad
plot(clicked, income, xlab="Income", ylab="Clicked on ad")

```
```{r}
# Age vs time spent on site
plot(site_time, age, xlab="Time spent on site", ylab="Age")
```

## Implementing the solution
```{r}
#Create a dataframe that selects those that clicked on an ad
yes <-  adverts %>% filter(adverts$clicked.on.ad  == 1); 
```
```{r}
# A summary of those that clicked
summary(yes)

```

```{r}
#See the various distributions

#Age distribution
hist(yes$age, col='maroon')

#See the area_income distribution 
hist(yes$area.income, col='yellow')

#See the daily_internet_usage distribution 
hist(yes$daily.internet.usage, col='blueviolet')

#daily_time_spent_on_site distribution 
hist(yes$daily.time.spent.on.site, col='grey')
```
```{r}
#Distribution of the countries 

countries <- table(yes$country)
countries
```
# Modeling
```{r}
names(data_num)
```

## SVM
# Splitting the data into training set and testing set. 
install.packages("numDeriv")

```{r}
library(numDeriv)
library(caret)
intrain <- createDataPartition(y = data_num$clicked.on.ad, p= 0.7, list = FALSE)
training <- data_num[intrain,]
testing <- data_num[-intrain,]
```

```{r}
# We check the dimensions of out training dataframe and testing dataframe
dim(training); 
dim(testing)
```
# Factorizing the categorical variable
```{r}
training[["clicked.on.ad"]] = factor(training[["clicked.on.ad"]])
```
# Controlling all the computational overheads using the traincolntrol() method
```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

```
# Checking the result of the train model
```{r}
svm_Linear <- train(clicked.on.ad ~., data = training, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneLength = 10)

```
# Making Predictions
```{r}
test_pred <- predict(svm_Linear, newdata = testing)
test_pred
```
# Plotting a confusion matrix to check for accuracy
```{r}
confusionMatrix(table(test_pred, testing$clicked.on.ad))
```
##
```{r}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

```{r}

# a) linear algorithms
set.seed(7)
fit.lda <- train(clicked.on.ad~., data=training, method="lda", metric=metric, trControl=control)
# b) nonlinear algorithms
# CART
set.seed(7)
fit.cart <- train(clicked.on.ad~., data=training, method="rpart", metric=metric, trControl=control)
# kNN
set.seed(7)
fit.knn <- train(clicked.on.ad~., data=training, method="knn", metric=metric, trControl=control)
# c) advanced algorithms
# SVM
set.seed(7)
fit.svm <- train(clicked.on.ad~., data=training, method="svmRadial", metric=metric, trControl=control)
# Random Forest
set.seed(7)
fit.rf <- train(clicked.on.ad~., data=training, method="rf", metric=metric, trControl=control)

```

# 
```{r}
# summarize accuracy of models
results <- resamples(list(lda=fit.lda, cart=fit.cart, knn=fit.knn, svm=fit.svm, rf=fit.rf))
summary(results)
```
```{r}
# compare accuracy of models
dotplot(results)
```

## Summarizing the best model
```{r}
print(fit.svm)
```

# Making Predictions
```{r}
# estimate skill of SVM on the training dataset
predictions <- predict(fit.svm, training)
confusionMatrix(predictions, training$clicked.on.ad)
```
# Conclusions
Support Vector Machine (SVM) is so far the  best model with an accuracy score of 96.86%. It made 678 correct predictions and 22 wrongs ones. The model can be further improved by tweaking some of the parameters. 